---
title: "Data simulation for distributed interpolation"
author: "Lauri Lovén"
date: "5/7/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.path = 'img/',
                      fig.width = 5,
                      fig.height = 4,
                      cache = TRUE,
                      autodep = TRUE)
library(gstat)
library(sp)
library(ggplot2)
library(patchwork)
library(viridis)
library(tidyverse)
theme_set(theme_minimal())

library
```

# Data

We compose the temperature field frames out of two components:

  1. the ambient weather conditions, following a smooth, slowly varying 2-dimensional Matern variogram model; and
  1. a 2-dimensional Gaussian variogram with short and long distance autocorrelation, simulating the surface contour map as
     well as long distance autocorrelation such as traffic, subway lines etc.

We set the following as constants.

```{r const, echo=TRUE}
# grid
x_max <- 100 # x coordinates range from 1 to x_max
y_max <- 100 # y coordinates range from 1 to y_max
x <- 1:x_max # x coordinates
y <- 1:y_max # y coordinates

# variogram hyperparameters
dist_long <- 100
dist_short <- 10
static_intercept <- 15
weather_intercept <- 0

##
##  HOX! Adjust n_frames for more or less training data.
##
n_frames <- 100                # data frames
observations_per_frame <- 100  # number of sensor observations for each frame
```


## Ambient weather conditions

Simulate the ambient weather conditions with a zero-centred Matern variogram. 

```{r weather}

w_dat <- expand.grid(x = x, y = y) # create data frame with all combinations

w_dat$z <- 1 # initialize z variable
coordinates(w_dat) <- ~x + y # set coordinates
gridded(w_dat) <- TRUE # specify data is gridded

g <- gstat(
  id= 'z',
  formula = z ~ 1,
  model = vgm(
    psill = 1,
    #psill = 10,
    #model = 'Sph',
    model = 'Mat',
    range = dist_long,
    kappa = 1,
    #kappa = 0.1,
    #anis = c(45 ,0.1)
    anis = c(90 , 0.9)
  ),
  data = w_dat,
  dummy = TRUE,
  beta = weather_intercept,
  maxdist = dist_short,
  nmax = 10
) # create gstat object

set.seed(123)
w_dat <- data.frame(predict(g, newdata = w_dat, nsim = n_frames)) # simulate random field data. 

```

Plot three different weather conditions.

```{r plot_weather_sample, include = TRUE, fig.width=6, fig.height=2}
sim_w1 <- ggplot(
  w_dat %>% rename(Temperature = sim1),
  aes(
    x = x,
    y = y,
    fill = Temperature)
  ) + 
  geom_raster() + 
  scale_fill_gradientn(colours = rainbow(7), limits = c(-4,4)) +
  labs(x = "", y = "") +
  theme(legend.position="none") +
  ggtitle("Weather frame 1")

sim_w2 <- ggplot(
  w_dat %>% rename(Temperature = sim2),
  aes(
    x = x,
    y = y,
    fill = Temperature)
  ) + 
  geom_raster() + 
  scale_fill_gradientn(colours = rainbow(7), limits = c(-4,4)) +
  labs(x = "", y = "") +
  theme(legend.position="none") +
  ggtitle("Weather frame 2")

sim_w3 <- ggplot(
  w_dat %>% rename(Temperature = sim3),
  aes(
    x = x,
    y = y,
    fill = Temperature)
  ) + 
  geom_raster() + 
  scale_fill_gradientn(colours = rainbow(7), limits = c(-4,4)) +
  labs(x = "", y = "") +
  theme(legend.position="none") +
  ggtitle("Weather frame 3")

sim_w1 + sim_w2 + sim_w3

```

## Spatial autocorrelation map

Simulate the contour map with an exp variogram with intercept 15. 

```{r contour}
c_dat <- expand.grid(x = x, y = y) # create data frame with all combinations
c_dat$z <- 1 # initialize z variable
coordinates(c_dat) <- ~x + y # set coordinates
gridded(c_dat) <- TRUE # specify data is gridded

g_c <- gstat(
  id= 'z',
  formula = z ~ 1 + y,
  #locations = ~x + y,
  model = vgm(
    #psill = 0.9,
    psill = 1,
    model = 'Exp',
    #model = 'Mat',
    range = dist_short,
    kappa = 1,
    #kappa = 0.1,
    #anis = c(45 ,0.1)
    anis = c(45 , 0.5)
  ),
  data = c_dat,
  dummy = TRUE,
  beta = c(static_intercept,0.03,0),
  maxdist = dist_short,
  nmax = 10
) # create gstat object

set.seed(456)
c_dat <- data.frame(predict(g_c, newdata = c_dat, nsim = 1)) # simulate random field data

ggplot(
  c_dat %>% rename(z = sim1),
  aes(
    x = x,
    y = y,
    fill = z)
  ) + 
  geom_raster() + 
  scale_fill_gradientn(colours = rainbow(7), limits = c(10, 21)) +
  labs(x = "", y = "", z = "") +
  theme(legend.position="none")
```


Finally, simulate long-distance covariance structures with an exp variogram. 

```{r long_autocov}
l_dat <- expand.grid(x = x, y = y) # create data frame with all combinations
l_dat$z <- 1 # initialize z variable
coordinates(l_dat) <- ~x + y # set coordinates
gridded(l_dat) <- TRUE # specify data is gridded

l_c <- gstat(
  id= 'z',
  formula = z ~ 1 + y,
  #locations = ~x + y,
  model = vgm(
    #psill = 0.9,
    psill = 1,
    model = 'Mat',
    #model = 'Mat',
    range = dist_long,
    kappa = 1
    #kappa = 0.1,
    #anis = c(45 ,0.1)
    #anis = c(45 , 0.5)
  ),
  data = l_dat,
  dummy = TRUE,
  #beta = c(1,0.03,0),
  beta = c(0),
  maxdist = dist_short,
  nmax = 10
) # create gstat object

set.seed(789)
l_dat <- data.frame(predict(l_c, newdata = l_dat, nsim = 1)) # simulate random field data
#names(dat)[3] <- 'z'
#head(dat) 

ggplot(
  l_dat %>% rename(z = sim1),
  aes(
    x = x,
    y = y,
    fill = z)
  ) + 
  geom_raster() + 
  scale_fill_gradientn(colours = rainbow(7), limits = c(-3,3)) +
  labs(x = "", y = "") +
  theme(legend.position="none")
```

## Compilation

The static element comprises the contour map and the long-distance autocorrelation. This is what we are trying to capture with our models.

```{r compiliation_static}
static_dat <- tibble(
  x = l_dat$x,   
  y = l_dat$y,
  z = l_dat$sim1 + c_dat$sim1
)

ggplot(
  static_dat,
  aes(
    x = x,
    y = y,
    fill = z)
  ) + 
  geom_raster() + 
  scale_fill_gradientn(colours = rainbow(7), limits = c(11, 23)) +
  labs(x = "", y = "") +
  theme(legend.position="none")
```

Weather data provides the perturbations to the static data set. Plot the first simulation and store the data.

```{r compiliation_final, fig.height=2, fig.width=6}
sim_dat <- as_tibble(lapply(w_dat[,-c(1,2)], function(x) x + static_dat$z))
sim_dat$x <- w_dat$x
sim_dat$y <- w_dat$y

sim1_plot <- ggplot(
  sim_dat %>% rename(z = sim1),
  aes(
    x = x,
    y = y,
    fill = z)
  ) + 
  geom_raster() + 
  scale_fill_gradientn(colours = rainbow(7), limits = c(11, 24)) +
  labs(x = "", y = "") +
  #theme(legend.position="none") +
  ggtitle("Simulation 1")

sim2_plot <- ggplot(
  sim_dat,
  aes(
    x = x,
    y = y,
    fill = sim2)
  ) + 
  geom_raster() + 
  scale_fill_gradientn(colours = rainbow(7), limits = c(11, 24)) +
  labs(x = "", y = "") +
  theme(legend.position="none") +
  ggtitle("Sim 2")


sim3_plot <- ggplot(
  sim_dat,
  aes(
    x = x,
    y = y,
    fill = sim3)
  ) + 
  geom_raster() + 
  scale_fill_gradientn(colours = rainbow(7), limits = c(11, 24)) +
  labs(x = "", y = "") +
  theme(legend.position="none")   +
  ggtitle("Sim 3")

sim1_plot + sim2_plot + sim3_plot + plot_layout(guides = 'collect')
```

Sample `r x_max` x coordinates and `r y_max` y coordinates for each of the `r n_frames` uniformly. 

```{r observations}

set.seed(123)
x_obs = lapply(1:n_frames, function(x) sample(1:x_max, observations_per_frame, replace = TRUE))

set.seed(456)
y_obs = lapply(1:n_frames, function(x) sample(1:x_max, observations_per_frame, replace = TRUE))

observations <- 
    lapply(
      1:n_frames,  
      function(n) {
        sim_frame <- sim_dat[[n]]
        obs_x <- x_obs[[n]] 
        obs_y <- y_obs[[n]]
        offsets <- (obs_y - 1) * x_max + obs_x
        return(tibble(x=obs_x, y=obs_y, z=sim_frame[offsets]))
      }
    )
  
```



Plot the first of the simulated observation sets.
```{r include=TRUE, fig.width=4, fig.height=6}
plot_obs <- function(n) {
  ggplot(
    observations[[n]],
    aes(
      x = x,
      y = y,
      fill = z)
    ) + 
    geom_raster() + 
    scale_fill_gradientn(colours = rainbow(7), limits = c(11, 24)) +
    labs(x = "", y = "") +
    theme(legend.position="none")   +
    ggtitle(paste("Observations ", n))
}

obs1_plot <- plot_obs(1)
obs2_plot <- plot_obs(2)
obs3_plot <- plot_obs(3)
(obs1_plot | sim1_plot) / (obs2_plot | sim2_plot) / (obs3_plot | sim3_plot) + plot_layout(guides = 'collect') 

```


Serialize data.

```{r serialize, echo=FALSE}
write_csv(sim_dat, "data_sim/sim_dat_truth.csv")
foo <- lapply(
  1:n_frames,
  function(n) {
    write_csv(observations[[n]], paste("data_sim/sim_dat_obs", n, ".csv", sep = ""))
  }
)
rm(foo)
print("Simulated truth and observations serialized.")     
```
